---
title: "ARIMA Model Guide: Tesla Revenue"
description:
author:
  - name: Mia Fusillo
    url: https://fusillomia.github.io/miaRBlog/
date: 5/31/2022
output:
  distill::distill_article:
    self_contained: false
---
### Packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(rvest)
library(fpp3) 
```

### Web Scraping
In my previous and first blog post, I manually scraped data off of the web. The data had less than 50 periods, so it wasn't too tedious to do by hand. However, this is clearly inefficient and would not be an option for larger datasets. For this blog, I read the HTML off of macro trends and used subsequent R syntax to scrape the specific data from the table on the page.

```{r}
URL <- "https://www.macrotrends.net/stocks/charts/TSLA/tesla/revenue"

stats <- URL %>% 
  read_html %>% 
  html_nodes("table") %>% 
  html_table()

revenue <- as.data.frame(stats[2])

colnames(revenue) = c("quarter", "revenue")
revenue$time <- seq.int(nrow(revenue))

revenue = revenue %>%
  arrange(-time)

revenue = revenue[4:52,]
revenue$time <- seq.int(nrow(revenue))

revenue <- revenue %>%
  add_column(qtr=yearquarter("2010 Q1") + 0:48, .before=TRUE) %>%
  as_tsibble(index=qtr)

revenue = revenue[,c(4,1,3)]

show(revenue)
```

### Data Cleaning
tsclean is a function that enables us to check for outliers and missing values.

### ARIMA Model Selection
The ARIMA command will filter through many different combinations of ARIMA models, specifically where p and q are respectively less than or equal to 3, and their sum is no more than 5.
